### 2018-08-17 doc/methods.tex
grew methods more

The logistic regression can thus be expressed as 
\begin{equation}
  \begin{aligned}
    t_{i} &\sim \text{Bernoulli}(\Theta) \\
    \Theta &= logit^{-1}(\beta_{w[i], f[i]} + \beta_{w[i], f[i]} r_{i} + \beta_{w[i], f[i]} d_{i} + \gamma_{1} c_{w[i]} + \gamma_{2} l_{w[i]} + a_{d[i], f[i]}) \\
    \beta_{w, f} &\sim MVN(\alpha_{f}, \Sigma) \\
    \alpha_{f} &\sim MVN(0, \Sigma) \\
    \gamma &\sim N(0, \sigma) \\
    a_{w, f} &\sim MVN(\delta_{f}, \Sigma) \\
    \delta_{f} &\sim MVN(0, \Sigma) \\
  \end{aligned}
  \label{eq:core}
\end{equation}
with \(i\) indexing the observation and bracket subscripts referencing the class of the \(i\)th observation e.g. \(j[i]\) is the age of the \(i\)th observation. \(a_{j[i]}\) is taxon age in bins at time of observation where \(j = 1, 2, \cdots, J\) and \(J\) is the maximum observed age of any taxon. \(b_{k}\) is the time of observation where \(k = 1, 2, \cdots, K\) and \(K\) is the number of time bins observed. Finally, \(c_{f}\) is the taxonomic group of an observation where \(f = 1, 2, \cdots, F)\) and \(F\) is the total number of taxonomic groups analyzed.

% priors

The basic core of the model described above \ref{eq:core} can be written in R formula syntax as
\begin{equation}
  \begin{aligned}
    \text{event} &\sim \text{range} + \text{range\_diff} + \text{temp} + \text{temp\_lag} + \\
    &\quad (1 + \text{range} + \text{range\_diff} | \text{mybin} / \text{phylum}) + \\
    &\quad (1 | \text{age} / \text{phylum}).
  \end{aligned}
\end{equation}
This formula is then used with a call to \begin{texttt} stan\_glmer \end{texttt} as Bernoulli family model which, along with the relevant data and prior choices, yields parameter estimates.

>>>>

The logistic regression can thus be expressed as 
\begin{equation}
  \begin{aligned}
    t_{i} &\sim \text{Bernoulli}(\Theta) \\
    \Theta_{i} &= \text{logit}^{-1} (X_{i} B_{w[i], p[i]} + a_{d[i], p[i]}) \\
    B_{w, p} &\sim MVN(\alpha_{p}, \Sigma_{B}) \\
    \alpha_{p} &\sim MVN(\mu, \Sigma_{\alpha}) \\
    a_{w, p} &\sim MVN(\delta_{p}, \Sigma_{a}) \\
    \delta_{p} &\sim MVN(0, \Sigma_{\delta}) \\
  \end{aligned}
  \label{eq:core}
\end{equation}
with \(i\) indexing the observation and bracket subscripts referencing the class of the \(i\)th observation where \(w[i]\) is the time of the \(i\)-th observation, \(p[i]\) is the phylum of the \(i\)-th observation, and \(d[i]\) is the age of the \(i\)-th observation.

\(B\) is a matrix of regression coefficients indexed by both time of observation and phylum.

\(\mu\) is a vector, the first element of which corresponds to the overall model intercept and the others being the overall average estimate of the regression coefficients.

To complete the generative model, we need to assign final priors to the ``top level'' parameters
\begin{equation}
  \begin{aligned}
    \mu_{c} &\sim 
      \begin{cases}
        N(0, 10) & \text{if } c = 1 \\
        N(0, 1) & \text{if } c > 1 \\
      \end{cases}
    \delta &\sim N(0, 1) \\
    \Sigma_{B} &= diag(\tau_{B}) \Omega_{B} diag(\tau_{B}) \\
    \Sigma_{\alpha} &= diag(\tau_{\alpha}) \Omega_{\alpha} diag(\tau_{\alpha}) \\
    \Sigma_{a} &= diag(\tau_{a}) \Omega_{a} diag(\tau_{a}) \\
    \Sigma_{\delta} &= diag(\tau_{\delta}) \Omega_{\delta} diag(\tau_{\delta}) \\
    \tau_{B} &\sim C^{+}(5) \\
    \tau_{\alpha} &\sim C^{+}(5) \\
    \tau_{a} &\sim C^{+}(5) \\
    \tau_{\delta} &\sim C^{+}(5) \\
    \Omega_{B} &\sim LKJ(1) \\
    \Omega_{\alpha} &\sim LKJ(1) \\
    \Omega_{a} &\sim LKJ(1) \\
    \Omega_{\delta} &\sim LKJ(1). \\
  \end{aligned}
  \label{eq:priors}
\end{equation}
The decomposition of the covariance matrices (e.g \(\Sigma_{B}\)) allows priors the variance and correlation aspects of covariance to be defined separately. These priors are also slightly more interpretable than a, for example, inverse-Wishart distributed prior for the covariance matrices. This decomposition is recommended by CITATION.

The basic core of the model described above \ref{eq:core} can be written in R formula syntax as
\begin{equation}
  \begin{aligned}
    \text{event} &\sim \text{range} + \text{range\_diff} + \text{temp} + \text{temp\_lag} + \\
    &\quad (1 + \text{range} + \text{range\_diff} | \text{mybin} / \text{phylum}) + \\
    &\quad (1 | \text{age} / \text{phylum}).
  \end{aligned}
\end{equation}
This formula is then used with a call to \begin{texttt} stan\_glmer \end{texttt} as Bernoulli family model which, along with the relevant data and prior choices, yields parameter estimates. An example line of R code to fit this mode given a data.frame ``data'' could be written as:
\begin{verbatim}
  stan_glmer(formula = event ~ range + range_diff + temp + temp_lag + 
                       (1 + range + range_diff | mybin/phylum) + 
                       (1 | age/phylum), 
             prior = normal(0, 1, autoscale = FALSE), 
             prior_intercept = normal(0, 10, autoscale = FALSE), 
             prior_aux = cauchy(0, 5, autoscale = FALSE), 
             data = data, 
             adapt_delta = 0.9999, 
             thin = 4)
\end{verbatim}

See our code repository LINK for full code details. Our code uses ``tidyverse'' tools such as dplyr CITATION, purr CITATION, and tidybayes CITATION, thus some familiarity with that package ecosystem is necessary to fully comprehend how we've processed our data and results.
